---
title: "Activity2"
author: "Gustavo T"
date: "2025-02-18"
output: html_document
---
This question involves the use of multiple linear regression on the Auto data set. The Auto data set are
available in the ISLR2 library.

Use the lm() function to perform a multiple linear regression with mpg as the response and all other
variables except name as the predictors. Use the summary() function to print the results
```{r}
library(ISLR2)
library(car)
attach(Auto)

summary(Auto)

autoModel1 <- lm(mpg~.-name, Auto)
autoModel1

summary(autoModel1)
```

Use the plot() function to produce diagnostic plots of the linear regression fit. Conduct appropriate tests.
Comment on any problems you see with the fit.
For example, comments on the following:
Linearity of the data.
- The data fairly supports a nonlinear model as we can see a parabola (quadratic) shape with a 
  wider funnel as we progress the x line.
Normality of residuals.
- As far as normality goes, the Q-Q plot seems to fit the line y = x but we start to notice more 
  outliers/high leverage points the higher you move up on the x-axis.
- Upon looking at the shapiro test, the p-value indicates that residual data is not normally
  distributed.
Homogeneity of residuals variance.
- From the ncvTest we see a small p-value at almost 0 meaning we lack the presence of a
  constant variance. The plot also does not indicate a horizontally flat red line, indicating
  there may not be constant variance.
Independence of residuals error terms.
- From the Durbin-Watson test we get a p-value of 0 indicating we have violated
  the independence assumption, therefore, residuals are not independent.
Do the residual plots suggest any unusually large outliers?
- 321, 323
Does the leverage plot identify any observations with unusually high leverage?
- 14
```{r}
plot(autoModel1, which = 1)
plot(autoModel1, which = 2)
plot(autoModel1, which = 3)
plot(autoModel1, which = 5)

shapiro.test(autoModel1$residuals)

ncvTest(autoModel1)

durbinWatsonTest(autoModel1)

rstudent(autoModel1) #studentized residuals
hatvalues(autoModel1) #leverage points
plot(predict(autoModel1), rstudent(autoModel1)) #plot the studentized residuals vs. the fitted values to detect outliers
plot(hatvalues(autoModel1), rstudent(autoModel1))#plot the studentized residuals vs. leverage to detect high leverage points
which.max(abs(rstudent(autoModel1))) #Gives the one that might be an outlier
which.max(hatvalues(autoModel1)) #Gives what might be a high-leverage point
hatvalues(autoModel1)>0.02
rstudent(autoModel1)>abs(3)
```

This question should be answered using the Carseats data set. The Carseats data set are available in the
ISLR2 library.
(a) Fit a multiple regression model to predict Sales using Price, Income, Advertising, and Population.
```{r}
attach(Carseats)
head(Carseats)

carseatModel <- lm(Sales~Price+Income+Advertising+Population, Carseats) # model with pop
carseatModel

summary(carseatModel)
```

(b) Provide an interpretation of each coefficient in the model.
  - Based on the coefficients obtained we can tell that with all features at 0
    we can expect 12.32599 units in sales and a decrease in sale number if there
    is an increase in price by 0.0539 units.;
  - an increase in income is associated with an increase in sales by 0.01098 units;
  - an increase in advertising will result in increased sales by 0.1238 units;
  - an increase in population indicates a lower number/value of sales by 0.0006 units.
```{r}
summary(carseatModel) # with population
```

(c) Write out the model in equation form.
  - Sales = -0.0538(Price) + 0.0110(Income) + 0.1202(Advertising) - 0.0006(Population)
  - Sales = -0.0538(Price) + 0.0110(Income) + 0.1202(Advertising)
  
(d) For which of the predictors can you reject the null hypothesis H0 : βj = 0?
  - We can reject the null hypothesis with Price, Income, and Advertising since they are 
    all significant predictors for Sales.

(e) On the basis of your response to the previous question, fit a smaller model that only uses the predictors
for which there is evidence of association with the outcome.
```{r}
carseatModel2 <- lm(Sales~Price+Income+Advertising, Carseats)
carseatModel2

summary(carseatModel2) # without population (better fit since pop. is not significant)
```

(f) How well do the models in (a) and (e) fit the data?
  - Considering both models have a p-value of basically 0, they both fit the data relatively well
    but after considering selected features and their level of significance, we can say
    that model 2 is the better one since it does not take non-significant features into account.
    It is also important to say that since both have a low R_squared they would not be considered
    to fit the data well.
```{r}
anova(carseatModel2, carseatModel)
```

(g) Using the model from (e), obtain 95 % confidence intervals for the coefficient(s).
```{r}
confint(carseatModel2, level = 0.95)
```

(h) Using the model from (e), produce diagnostic plots of the linear regression fit. Conduct appropriate tests.
Comment on any problems you see with the fit.
For example, comments on the following:
Linearity of the data.
- The second model has a much flatter horizontal line for residuals vs fitted
  giving it a much higher linearity coefficient. The points are also scattered
  randomly not forming any type of pattern.
Normality of residuals.
- While looking at the Q-Q plot we can see the vast majority of points
  lie on the line y = x, indicating normality, with outliers/high leverage
  points only being located at both tail ends. In addition, the shapiro test
  outputs a p-value of 0.4779 which indicates normality since > 0.05.
Homogeneity of residuals variance.
- The red line for the plot is mostly flat indicating the possibility of
  constant variance. Through the ncvTest we get a p-value of 0.3519, which
  is higher than 0.05 showing homogeneity, or evidence of constant variance.
Independence of residuals error terms.
- Looking at the residual vs leverage point plot we can see that the majority
  of points are located around 0 with no clear pattern which can indicate independence.
  Furthermore, the Durbin-Watson test shows a p-value of 0.474 which is higher than 0.05
  and means there is no significant autocorrelation (independence acquired). 
Do the residual plots suggest any unusually large outliers?
- 51
Does the leverage plot identify any observations with unusually high leverage?
- 43
```{r}
plot(carseatModel2, which = 1)
plot(carseatModel2, which = 2)
plot(carseatModel2, which = 3)
plot(carseatModel2, which = 5)

shapiro.test(carseatModel2$residuals)

ncvTest(carseatModel2)

durbinWatsonTest(carseatModel2)

which.max(abs(rstudent(carseatModel2))) #Gives the one that might be an outlier
which.max(hatvalues(carseatModel2)) #Gives what might be a high-leverage point
```

