---
title: "Class Activity 7"
output: html_document
date: "2025-04-22"
---
In this exercise, you will use Boston data set to perform best subset selection, where per capita crime rate is the response variable and all other remaining variables as predictors. Note: The Boston data set is in the MASS library
```{r}
library(MASS)
library(ISLR2)
library(leaps)
attach(Boston)
```

(a)	Apply regsubsets() to the data set in order to perform best subset selection. What is the best model obtained according to Cp, BIC, and adjusted R^2? Show some plots to provide evidence for your answer, and report the coefficients of the best model obtained.
  - Adjusted R^2 -> 9 predictors
  - Cp -> 7 predictors
  - BIC -> 2 predictors
```{r}
dim(Boston)

regfit_full <- regsubsets(crim ~ ., data = Boston, nvmax = 12)
reg_summary <- summary(regfit_full)
reg_summary

par(mfrow=c(2,2))

plot(reg_summary$rss,xlab="Number of Variables",ylab="RSS")

which.max(reg_summary$adjr2)
plot(reg_summary$adjr2,xlab="Number of Variables",ylab="Adjusted RSq")
points(9,reg_summary$adjr2[9], col="red",cex=2,pch=20)
coef(regfit_full, 9)

which.min(reg_summary$cp)
plot(reg_summary$cp,xlab="Number of Variables",ylab="Cp")
points(7,reg_summary$cp[7],col="red",cex=2,pch=20)
coef(regfit_full, 7)

which.min(reg_summary$bic)
plot(reg_summary$bic,xlab="Number of Variables",ylab="BIC")
points(2,reg_summary$bic[2],col="red",cex=2,pch=20)
coef(regfit_full, 2)
```

(b)	Repeat (a), using forward stepwise selection and also using backwards stepwise selection. How does your answer compare to the results in (a)?
  - For adjusted R^2 value, both forward and backward selected 11 predictors while best subset selection
    opted for 9. Forward and backward selection may be over fitting the data compared to best subset
    selection which chose a simpler model which is able to explain the data similarly to the stepwise
    methods.
  - For Cp value, both forward and backward selected 10 while best subset selection chose 7. In this case,
    the stepwise methods decided to go for more complex models as opposed to best subset selection, 
    since they may not consider all combination like best subset selection.
  - For BIC value, both forward and best subset selection chose 2, while backward chose 8. lastly, 
    Backwards may not be reducing the size of the model as aggressively considering it starts with all
    predictors available and reduces them one at a time.
```{r}
# Forward
regfit_fwd=regsubsets(Salary~.,data=Hitters,nvmax=12,method="forward")
fwd_summary <- summary(regfit_fwd)
fwd_summary

par(mfrow=c(2,2))

plot(fwd_summary$rss,xlab="Number of Variables",ylab="RSS")

which.max(fwd_summary$adjr2)
plot(fwd_summary$adjr2,xlab="Number of Variables",ylab="Adjusted RSq")
points(11,fwd_summary$adjr2[11], col="red",cex=2,pch=20)
coef(regfit_fwd, 11)

which.min(fwd_summary$cp)
plot(fwd_summary$cp,xlab="Number of Variables",ylab="Cp")
points(10,fwd_summary$cp[10],col="red",cex=2,pch=20)
coef(regfit_fwd, 10)

which.min(reg_summary$bic)
plot(fwd_summary$bic,xlab="Number of Variables",ylab="BIC")
points(2,fwd_summary$bic[2],col="red",cex=2,pch=20)
coef(regfit_fwd, 2)
```

```{r}
# Backwards
regfit_bwd=regsubsets(Salary~.,data=Hitters,nvmax=12,method="backward")
bwd_summary <- summary(regfit_bwd)
fwd_summary

par(mfrow=c(2,2))

plot(bwd_summary$rss,xlab="Number of Variables",ylab="RSS")

which.max(bwd_summary$adjr2)
plot(bwd_summary$adjr2,xlab="Number of Variables",ylab="Adjusted RSq")
points(11,bwd_summary$adjr2[11], col="red",cex=2,pch=20)
coef(regfit_bwd, 11)

which.min(bwd_summary$cp)
plot(bwd_summary$cp,xlab="Number of Variables",ylab="Cp")
points(10,bwd_summary$cp[10],col="red",cex=2,pch=20)
coef(regfit_bwd, 10)

which.min(bwd_summary$bic)
plot(bwd_summary$bic,xlab="Number of Variables",ylab="BIC")
points(8,bwd_summary$bic[8],col="red",cex=2,pch=20)
coef(regfit_bwd, 8)
```

